Detailed Chemical Abundances with \tc{}

### 2015-10-27 article/ms.tex
Things

@@

\begin{equation}
\mathcal{W}_n = \sigma{}\sqrt{2\pi}(1 - f_{n\lambda_r}).
\label{eq:ew-flux}
\end{equation}

Here $\sigma$ is the standard deviation of the profile. For typical FGK-type
stars observed with $\mathcal{R} \lesssim 20,000$, $\sigma$ will be dominated
by the line spread function (LSF) of the spectrograph. Because \tc{} is best suited
to large spectroscopic surveys, we can presume that the LSF is well-known, or
at least reasonably constant for the training set, perhaps approximated by the 
quoted spectral resolution $\mathcal{R}$. In Section \ref{sec:something} we use
the knowledge of the spectrograph LSF, but for explanatory purposes here we will
assume it is constant for all $n$ stars surrounding the profile at $\lambda_r$.

@@

\noindent{}where the label vector 
$\lv = \{T_{\rm eff},\log{g},\log_{\epsilon}{\rm Si}\}$\footnote{Recall
${\rm [Si/Fe]} \equiv {\rm [Si/H]} - {\rm [Fe/H]} \equiv \log_{\epsilon}{\rm Si} 
- \log_{\epsilon}{\rm Si}_\odot - {\rm [Fe/H]}$, where we adopt the \citet{Asplund2009}
solar abundances.} and the coefficients of the vector $\psiv$ for a given 
atomic line need to be solved for. Combining Equations \ref{eq:ew-flux} and
\ref{eq:line-strength-model} yields the approximate flux $f_{n,\lambda_r}$ that 
atomic line $m$ contributes to star $n$ at wavelength $\lambda_r$, which we
describe as $h(\lv|\psiv,\sigma)$:

\begin{equation}
f_{n,\lambda_r} \approx h(\lv|\psiv,\sigma,\lambda) \equiv 1 - \frac{\psiv^{\intercal}\cdot\lv}{\sigma\sqrt{2\pi}}
\end{equation}

We calculate $\psiv$ given the stellar parameters of the training set and the
calculated $\logeps{(Si~1)}$ abundances from MOOG, calculated at many \ew\ values.
The difference between the MOOG-calculated abundance and our line strength model
is shown in the second panel of Figure \ref{fig:atomic-line-model-example}.
Negligible bias is present, and the standard deviation is 0.01~dex, demonstrating
that a quadratic-in-labels model can relate line strength \ew\ and abundance
for all types of stars in the training set. This is further shown in the third panel of 
Figure \ref{fig:atomic-line-model-example},
which shows the spectrum surrounding the Si 1 line for a random star\footnote{The
star is truly chosen at random, see Appendix \ref{appendix:reproducibility} for steps
to reproduce this research.} in the training set. We calculated the corresponding
line strength \ew\ given the stellar parameters and reported Si 1 abundance, 
adopted a spectral resolution of $\mathcal{R} = 22,500$ and produced a model 
spectrum of this atomic line following Equation \ref{ew-flux}.

We note that some of the 0.01~dex scatter between our line strength model and 
MOOG (second panel of Figure \ref{fig:atomic-line-model-example}) is likely due to 
the tolerance employed by MOOG's optimisation algorithm to calculate a line 
abundance, given an EW.

@@

% h relates W to flux values.
% W

% Discuss using the Cannon model from Ness et al to model the EW as a func. of stellar params.

@@

Knowing that a simple model can relate the a single line abundance 
$\log_{\epsilon}{\rm X}$ to the equivalent width \ew\, given the stellar
parameters, it is intriguing to ask whether model atmosphere relations and
the theory of spectral line formation can guide the level of model complexity
required.

@@

We now have a functional form of a model that relates $\mathcal{W}$ to $\log_\epsilon\left({\textrm X}\right)$ (or a scaled-solar [X/H] value), given the stellar parameters, which themselves are part of the set of $K$ labels. It is interesting to note that Equation \ref{eq:ew-logx-model} is not dissimilar to the quadratic-in-labels model used by \citet{Ness15a} for APOGEE data. With only the prior knowledge that \textit{a simple model of the labels can predict spectral fluxes}, we have derived the appropriate functional form for a single absorption line. The model $\mathcal{W}\left({\rm [X/H]}, T, logg\right)$ simply approximates the curve-of-growth for a single line, and the coefficients still need to be calculated. The practicalities of determining these coefficients is outlined in Section \ref{sec:somewhere}. For the purposes of introducing the model, we can assume that they are easily calculable.

Consider multiple atomic lines, where the coefficients to translate $\mathcal{W}$ and [X/H] are known. Given some knowledge of the line spread function, the EWs $\mathcal{W}$ to flux values $f_{n\lambda}$ can be translated such that:

\begin{equation}
h = X
\label{eq:h-model}
\end{equation}

Here we implicitly assume the mutual opacity contributions of nearby weak lines will be small. 
At this point it is worthwhile to ask whether Equation \ref{eq:h-model} can be scaled sufficiently to $M$ atomic lines in order to produce an entire spectral model for $f_{n\lambda}$, without the contribution of $g\left(\lv\given\cv\right)$. There would be a number of immediate issues with that approach: a huge number of additional coefficients would be required, hidden lines could not be encapsulated accurately by the model, strong lines would be extremely poorly modelled by the fit, no longer linear (slow), but most importantly the pixels would lose flexibility in training on arbitrary labels.

If lines are blended, one could consider modelling many more atomic transitions.
However as the number grows, this situation becomes unnecessarily complex:
often blended lines exist and we do not care about their abundance, only their
inclusion (nuisance or dirty lines). Other features may be a complex combination
of thousands of molecular features.

Here we introduce a method of including individual atomic transitions within an existing \textit{Cannon} model.

The spectral model becomes a multiplicative function of $g$ and $h$ with some
unknown noise.

@@

This inclusion combines the existing flexibility in the model, allowing for training on arbitrary labels, and uses individual absorption lines for which some information is known (species, atomic data, etc) in order to trace the chemical abundances of individual elements.

The vector $\wv$ can be calculated directly from spectral synthesis, or using data.
In practice, given some rest-frame normalised spectra and some knowledge of the
line spread function, $h(\lv,\wv\given{}X)$ can be calculated before training \tc.

%Utilising a simple polynomial model to predict spectral fluxes is a distinctive
%method from parallel efforts in radiative transfer of realistic (3D) stellar 
%photospheres. Calculations of this nature are extremely expensive. Indeed, even
%with simplifications of these models (e.g., interpolated 1D photospheres),
%calculating synthetic spectra requires knowledge of species opacities, detailed 
%atomic and molecular information for all contributing elements, solving for
%the equation of state, and iteratively performing expensive radiative transfer
%calculations. Given the
%simplicity of a quadratic-in-labels spectral model with \tc{}, it is remarkable
%that these complex calculations can be simplified so greatly, whilst maintaining
%the typical measurement fidelity reported for high-resolution spectroscopic studies based on
%\textit{curve-of-growth} analyses.

% Figure of relating EW to [X/H] with some simple model compared to analytic value.

The compromise is to use TC and `clean' lines that we care about in concert.

% existing simple model. replaces expensive synthesis, involving model atmosphere
% generation, t-tau relations, opacities, equation of state, integrated line-of-sight
% and radiative transfer, etc.

% before extending this model to include individual abundances, it is worthwhile
% demonstrating (from a physics perspective) *why* the cannon works so well.

% APOGEE, R of 20,000, 
% LAMOST, higher Rs
% super high-resolution also works.

@@

@@

>>>>

\long\def\/*#1*/{}

@@

\begin{equation}
\mathcal{W}_n = \sigma{}a\sqrt{2\pi}.
\label{eq:ew-flux}
\end{equation}

Here $a$ is the amplitude of the profile ($a = 1 - f_{n\lambda_r}$) and $\sigma$ 
is the standard deviation. For typical FGK-type
stars observed with $\mathcal{R} \lesssim 20,000$, $\sigma$ will be dominated
by the line spread function (LSF) of the spectrograph. Because \tc{} is best suited
to large spectroscopic surveys, we can presume that the LSF is well-known, or
at least reasonably constant for the training set, perhaps approximated by the 
quoted spectral resolution $\mathcal{R}$. In Section \ref{sec:something} we use
the knowledge of the spectrograph LSF, but for explanatory purposes here we will
assume it is constant for all $n$ stars surrounding the profile at $\lambda_r$.

@@

\noindent{}where the label vector 
$\lv = \{T_{\rm eff},\log{g},\log_{\epsilon}{\rm Si}\}$\footnote{Recall
${\rm [Si/Fe]} \equiv {\rm [Si/H]} - {\rm [Fe/H]} \equiv \log_{\epsilon}{\rm Si} 
- \log_{\epsilon}{\rm Si}_\odot - {\rm [Fe/H]}$, where we adopt the \citet{Asplund2009}
solar abundances.} and the coefficients of the vector $\psiv$ for a given 
atomic line need to be solved for. Thus from Equations \ref{eq:ew-flux} and
\ref{eq:line-strength-model} we obtain the approximate flux $f_{n,\lambda_r}$ that 
atomic line $m$ (at wavelength $\lambda_m$) contributes to star $n$ at any
wavelength $\lambda$:

\begin{equation}
f_{n,\lambda_r} \approx 1 - \left(\psiv^{\intercal}\cdot\lv\right)\mathcal{N}\left(\lambda_{m} - \lambda_r,\sigma\right).
\end{equation}

We calculate $\psiv$ given the stellar parameters of the training set and the
calculated $\logeps{(Si~1)}$ abundances from MOOG, calculated for multiple \ew.
The difference between the MOOG-calculated abundance and our line strength model
is shown in the second panel of Figure \ref{fig:atomic-line-model-example}.
The agreement is excellent. Negligible bias is present, and the standard deviation is 0.01~dex, demonstrating
that a quadratic-in-labels model can relate line strength \ew\ and abundance
for all types of stars in the training set. This is further shown in the third panel of 
Figure \ref{fig:atomic-line-model-example},
which shows the spectrum surrounding the Si 1 line for a random star\footnote{The
star is truly chosen at random, see Appendix \ref{appendix:reproducibility} for steps
to reproduce this research.} in the training set. We calculated the corresponding
line strength \ew\ given the stellar parameters and reported Si 1 abundance, 
adopted a spectral resolution of $\mathcal{R} = 22,500$ and produced a model 
spectrum of this atomic line following Equation \ref{ew-flux}. We note that some 
of the 0.01~dex scatter between our 
line strength model and MOOG (second panel of Figure 
\ref{fig:atomic-line-model-example}) is likely due to the tolerance employed by 
MOOG's optimisation algorithm to calculate a line abundance, given an EW.

@@

@@

Knowing that a quadratic model can relate the a single line abundance 
$\logeps{X}$ to the equivalent width \ew\, given the stellar
parameters, it is intriguing to ask whether model atmosphere relations and
the theory of spectral line formation can guide the level of model complexity
required.

@@

We now have a physically-motivated model that relates \ew\ to $\logeps{X}$
(or a scaled-solar [X/H] or [X/Fe] ratio), given the stellar parameters. Here we
describe how we include multiple atomic line models as part of \tc's training
and testing phases.

% different psi_m for each line at lamgda_m

% Introduce a h function that includes treatment from all lines.

\begin{equation}
h(\lv\given\{\psiv\},\{\lambda_m\},\sigma) = 1 - \prod^{M}_{m=0}(\psiv^\intercal\cdot\lv)\mathcal{N}(\lambda_m - \lambda,\sigma)
\end{equation}

% Assumptions for the training set (complete abundances). Incomplete
% abundances in the training set can be discussed in Section ref.

@@

The training stage can be described in practice as the following:

Build all the atomic lines of interest.

For each star
\begin{enumerate}
\item Use the reference abundance for each elemental species to calculate expected EWs
\item Use the LSF to calculate $h$
\item Divide the flux by $h$
\end{enumerate}

For each pixel
\begin{enumerate}
\item Train \tc{} on the residuals
\end{enumerate}

This framework combines the existing flexibility in \tc\, allowing for training
on arbitrary labels, and uses individual absorption lines for which some 
information is known (species, atomic data, etc) in order to trace the chemical 
abundances of individual elements.

At testing time, the story is a little different. Because the stellar parameters
influence both $h$ and $g$, we must solve for the stellar parameters and all
chemical abundances simultaneously.

@@

We will use the APOGEE X sample to derive individual chemical abundances of 15
elements using \tc{}.

@@

\/* 
% for the discussion.
Here we implicitly assume the mutual opacity contributions of nearby weak lines will be small. 
At this point it is worthwhile to ask whether Equation \ref{eq:h-model} can be scaled sufficiently to $M$ atomic lines in order to produce an entire spectral model for $f_{n\lambda}$, without the contribution of $g\left(\lv\given\cv\right)$. There would be a number of immediate issues with that approach: a huge number of additional coefficients would be required, hidden lines could not be encapsulated accurately by the model, strong lines would be extremely poorly modelled by the fit, no longer linear (slow), but most importantly the pixels would lose flexibility in training on arbitrary labels.

If lines are blended, one could consider modelling many more atomic transitions.
However as the number grows, this situation becomes unnecessarily complex:
often blended lines exist and we do not care about their abundance, only their
inclusion (nuisance or dirty lines). Other features may be a complex combination
of thousands of molecular features.
*/
